---
title: "My Project"
author: "JS"
date: "`r Sys.Date()`"
output: html_document
---

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r fig.width=20}

# library(pacman)
# 
# pacman::p_load(
#   ggplot2,
#   sf,
#   raster,
#   rnaturalearth,
#   rnaturalearthdata,
#   gadm,
#   here,
#   terra,
#   tidyr,
#   tidyverse,
#   dplyr,
#   magrittr,
#   tidyterra,
#   rnaturalearthhires
# )
# 
# here::i_am("exercises_code/My Project.Rmd")
# 
# wine = rast(here("reben_2010.tiff"))
# names(wine) = c("r", "g", "b")
# wine %<>%
#   filter(!(r == 255 & g == 255 & b == 255))
# 
# basemap_sf = ne_states(country = "Germany", returnclass = "sf") %>% 
#   select(1) %>% 
#   st_transform(crs = crs(wine))
# 
# # basemap_sf = ne_countries(country = "Germany", 
# #                           scale = "medium", 
# #                           returnclass = "sf") %>% 
# #   dplyr::select(1) %>% 
# #   st_transform(crs = crs(wine))
# 
# basemap_vect = vect(basemap_sf)
# 
# plot(basemap_vect)
# plot(wine, alpha = 1, add = T, legend=T, axes=T)
```

```{r}
# wms_saxony_dgm_url <- "https://geodienste.sachsen.de/wms_geosn_hoehe/guest?VERSION=1.3.0"
# wms_saxony_wine_url <- "https://www.smul.sachsen.de/gis-online/services/Weinbau_public/MapServer/WmsServer"
# wms_saxony_terrain_url <- "https://luis.sachsen.de/arcgis/services/boden/bk25wein/MapServer/WMSServer"
# 
# ginfo_dgm <- sf::gdal_utils('info', str_c("WMS:", wms_saxony_dgm_url), quiet = T) 
# lurl_dgm <- ginfo_dgm %>% str_match_all('SUBDATASET_(\\d)_NAME=(.*?)\n') %>% .[[1]] %>% .[, 3]
# ldesc_dgm = ginfo_dgm %>% str_match_all("SUBDATASET_(\\d)_DESC=(.*?)\n") %>% .[[1]] %>% .[, 3]
# wms_layers_dgm <- cbind(ldesc_dgm, lurl_dgm)
# 
# saxony_map <- st_read(here("data","raw", "gadm41_DEU_shp", "gadm41_DEU_4.shp"), quiet = T) %>% 
#   filter(NAME_1 == "Sachsen") %>% 
#   dplyr::select(3)
# 
# plot(saxony_map)

```

```{r load packages}
pacman::p_load(
  sf, 
  httr,
  tidyverse,
  ows4R,
  here,
  gridExtra,
  terra
)
```

```{r Modification/Fix of OWSHttpRequest$private_methods$GET }

OWSHttpRequest$private_methods$GET = function (url, request, namedParams, mimeType) 
{
    namedParams <- c(namedParams, request = request)
    params <- paste(names(namedParams), namedParams, sep = "=", collapse = "&")
    req <- url
    
    
    if(nzchar(params)){
      if(!endsWith(url,"?")){
        if(regexpr("\\?", url)>0 & regexpr("/cas?service=", url, fixed = T)<0){
          req <- paste0(req, "&")
        }else{
          req <- paste0(req, "?")
        }
      }
    }
    if(regexpr("/cas?service=", url, fixed = T) > 0) params <- URLencode(params, reserved = TRUE)
    req <- paste0(req, params)
    self$INFO(sprintf("Fetching %s", req))
    
    headers <- private$headers
    if (!is.null(private$token)) {
        headers <- c(headers, Authorization = paste(private$auth_scheme, 
            private$token))
    }
    r <- NULL
    if (self$verbose.debug) {
        r <- with_verbose(GET(req, add_headers(headers), set_config(private$config)), 
            progress())
    }
    else if (self$verbose.info) {
        r <- GET(req, add_headers(headers), set_config(private$config), 
            progress())
    }
    else {
        r <- GET(req, add_headers(headers), set_config(private$config))
    }
    responseContent <- NULL
    if (is.null(mimeType)) {
        responseContent <- content(r, encoding = "UTF-8")
    }
    else {
        if (regexpr("xml", mimeType) > 0) {
            text <- content(r, type = "text", encoding = "UTF-8")
            text <- gsub("<!--.*?-->", "", text)
            responseContent <- xmlParse(text)
        }
        else if (regexpr("json", mimeType) > 0) {
            responseContent <- content(r, type = "text", encoding = "UTF-8")
        }
        else {
            responseContent <- content(r, type = mimeType, encoding = "UTF-8")
        }
    }
    response <- list(request = request, requestHeaders = httr::headers(r), 
        status = status_code(r), response = responseContent)
    return(response)
}

```

```{r, fig.width=10}

# wfs_laermindex <- "https://kommisdd.dresden.de/net3/public/ogcsl.ashx?nodeid=1047"
# 
# 
# laermindex_client = WFSClient$
#   new(wfs_laermindex, serviceVersion = "2.1.0")
# laermindex_feat_name = laermindex_client$getFeatureTypes(pretty = TRUE)[1, "name"]
# laermindex_client$getCapabilities()$findFeatureByName(laermindex_feat_name)
# 
# 
# url_laermindex <- parse_url(wfs_laermindex)
# url_laermindex$query <- c(url_laermindex$query, 
#                           list(service = "wfs", request = "GetFeature", 
#                                typenames = "regions", srsName = "EPSG:4326"))
# request <- build_url(url_laermindex)
# read_sf(request)


laermindex_sf = read_sf(here("data", "laermindex.gpkg"))
stadtteile_sf = read_sf(here("data", "stadtteile.gpkg"))

bevoelkerung_0_bis_5_jahre_sf = read_sf(here("data", "bevoelkerung_0_bis_5_jahre.gpkg"))
bevoelkerung_6_bis_13_jahre_sf = read_sf(here("data", "bevoelkerung_6_bis_13_jahre.gpkg"))
bevoelkerung_14_bis_17_jahre_sf = read_sf(here("data", "bevoelkerung_14_bis_17_jahre.gpkg"))
bevoelkerung_18_bis_26_jahre_sf = read_sf(here("data", "bevoelkerung_18_bis_26_jahre.gpkg"))
bevoelkerung_27_bis_59_jahre_sf = read_sf(here("data", "bevoelkerung_27_bis_59_jahre.gpkg"))
bevoelkerung_60_bis_74_jahre_sf = read_sf(here("data", "bevoelkerung_60_bis_74_jahre.gpkg"))
bevoelkerung_ab_75_jahre_sf = read_sf(here("data", "bevoelkerung_ab_75_jahre.gpkg"))

gesundheitseinrichtungen_sf = read_sf(here("data", "gesundheitseinrichtungen.gpkg"))
seniorenbegegnunsstaetten_sf = read_sf(here("data", "seniorenbegegnunsstaetten.gpkg"))
ambulante_mobile_dienste_sf = read_sf(here("data",  "ambulante_mobile_dienste.gpkg"))
betreutes_wohnen_sf = read_sf(here("data", "betreutes_wohnen.gpkg"))
senioren_und_pflegeheime_sf = read_sf(here("data", "senioren_und_pflegeheime.gpkg"))

dresden_stadtteile_centroids = st_centroid(stadtteile_sf)
dresden_centroid = st_centroid(st_union(stadtteile_sf))


ggplot() +
  geom_sf(data = stadtteile_sf) + geom_sf(data = dresden_centroid) + theme_minimal()

# saxony_map <- st_read(here("data","raw", "gadm41_DEU_shp", "gadm41_DEU_4.shp"), quiet = T) %>% 
#   filter(NAME_4 == "Dresden") %>%
#   dplyr::select(3)

# ggplot() + 
#   # geom_sf(data = saxony_map) + 
#   geom_sf(data = stadtteile_sf) +
#   # geom_sf(data = laermindex_sf) +
#   geom_sf(data = bevoelkerung_ab_75_jahre_sf, show.legend = T, aes(fill = prozent, color = prozent)) +
#   #geom_sf(data = gesundheitseinrichtungen_sf, col = "blue", size = 1) +
#   geom_sf(data = seniorenbegegnunsstaetten_sf, col = "orange", size = 1) +
#   #geom_sf(data = ambulante_mobile_dienste_sf, col = "yellow", size = 1) +
#   #geom_sf(data = betreutes_wohnen_sf, col = "green", size = 1) +
#   geom_sf(data = senioren_und_pflegeheime_sf, col = "red", size = 1)


plot_titles = c("0 bis 5 Jahre", "14 bis 17 Jahre", "18 bis 26 Jahre", 
           "27 bis 59 Jahre", "60 bis 74 Jahre", "Ab 75 Jahren")
names(plot_titles) = as.character(1:length(plot_titles))
do.call(grid.arrange, 
        c(imap(list(bevoelkerung_0_bis_5_jahre_sf, bevoelkerung_14_bis_17_jahre_sf,
                   bevoelkerung_18_bis_26_jahre_sf, bevoelkerung_27_bis_59_jahre_sf,
                   bevoelkerung_60_bis_74_jahre_sf, bevoelkerung_ab_75_jahre_sf), 
            \(sf_d, idx) ggplot() + 
              geom_sf(data = stadtteile_sf) + 
              geom_sf(data = sf_d, aes(fill = prozent, color = prozent)) +
              theme_minimal() + 
              ggtitle(plot_titles[idx])
              ), 
          ncol = 3, nrow = 2))

combined_bev_sf = dplyr::bind_rows(list(bevoelkerung_0_bis_5_jahre_sf, bevoelkerung_14_bis_17_jahre_sf,
                       bevoelkerung_18_bis_26_jahre_sf, bevoelkerung_27_bis_59_jahre_sf,
                       bevoelkerung_60_bis_74_jahre_sf, bevoelkerung_ab_75_jahre_sf), 
                       .id = "gruppe") 
combined_bev_sf$gruppe = plyr::revalue(as.factor(combined_bev_sf$gruppe), plot_titles)

ggplot() + 
  geom_line(data = combined_bev_sf, 
             mapping = aes(x = stadtteil, y = prozent, group = gruppe, col = gruppe)) +
  geom_point(data = combined_bev_sf, 
             mapping = aes(x = stadtteil, y = prozent, group = gruppe, col = gruppe)) +
  theme_bw() + 
  labs(title = "Population in Dresden by age groups and quarter, measured in 2021")

```

The biggest percentage of the population is always taken up by the people between 27 to 59 years, with one exception: in "Südvorstadt-Ost" (quarter id 82), the percentage of people between 18 and 26 years is slightly higher (37.4%) than the percentage of people between 27 and 59 years (35.2%). However, this is an outlier of the share of people between 18 and 26 years as their median percentage is 9.2% and their mean 10.84% with a standard deviation of 5.87%; the people between 27 and 59 years have the highest share with a mean of 43.63%, a median of 42.59% and a standard deviation of 5.13%.
