---
title: "Uni GIS Project"
author: "Julius Schmidt"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: true
    fig_caption: true
    number_sections: true
  html_document: default
papersize: a4
bibliography: [references.bib, packages.bib]
---

```{r setup, include = F, warning = F}
knitr::opts_chunk$set(warning = F, message = F)
# Install necessary dependencies from renv.lock
renv::restore()
packages = c(
  "knitr",
  "kableExtra",
  "sf", 
  "tidyverse",
  "here",
  "terra",
  "gridExtra",
  "cowplot",
  "reshape2",
  "magrittr",
  "tidygeocoder",
  "maptiles",
  "osmdata",
  "ncdf4",
  "exactextractr",
  "tidyterra",
  "osrm",
  "factoextra",
  "MASS",
  "FactoMineR"
)
lapply(packages, library, character.only = T)
#knitr::write_bib(file = 'packages.bib')
```

<!-- The report itself should be an R markdown or a Quarto R document of roughly 5 pages (counting without code but including meaningful images or tables that you create), and I should find inside of it a part where you explain the data structure and show summary statistics, a part where you present your research questions and your starting hypothesis, a part where you explain your chosen method of analysis and then apply it and finally a concluding part where you discuss what you find in relation to your hypothesis. -->

```{r Data Setup, echo=F, warning=F}
# Open Data Dresden Portal Data
stadtteile_sf = read_sf(here("data", "stadtteile.gpkg"))
bevoelkerung_0_bis_5_jahre_sf = read_sf(here("data", "bevoelkerung_0_bis_5_jahre.gpkg"))
bevoelkerung_6_bis_13_jahre_sf = read_sf(here("data", "bevoelkerung_6_bis_13_jahre.gpkg"))
bevoelkerung_14_bis_17_jahre_sf = read_sf(here("data", "bevoelkerung_14_bis_17_jahre.gpkg"))
bevoelkerung_18_bis_26_jahre_sf = read_sf(here("data", "bevoelkerung_18_bis_26_jahre.gpkg"))
bevoelkerung_27_bis_59_jahre_sf = read_sf(here("data", "bevoelkerung_27_bis_59_jahre.gpkg"))
bevoelkerung_60_bis_74_jahre_sf = read_sf(here("data", "bevoelkerung_60_bis_74_jahre.gpkg"))
bevoelkerung_ab_75_jahre_sf = read_sf(here("data", "bevoelkerung_ab_75_jahre.gpkg"))
ambulante_mobile_dienste_sf = read_sf(here("data",  "ambulante_mobile_dienste.gpkg"))
bahnhoefe_sf = read_sf(here("data", "bahnhoefe_dresden.gpkg"))
betreutes_wohnen_sf = read_sf(here("data", "betreutes_wohnen.gpkg"))
einwohnerzahlen_absolut = read_csv2(here("data", "einwohnerzahlen_absolut.csv"), col_names = F, show_col_types = F)
gesundheitseinrichtungen_sf = read_sf(here("data", "gesundheitseinrichtungen.gpkg"))
grundschulen_sf = read_sf(here("data", "grundschulen.gpkg"))
haltestellen_sf = read_sf(here("data", "haltestellen_dresden.gpkg"))
kindertagesstaetten_sf = read_sf(here("data", "kindertagesstaetten.gpkg"))
oberschulen_sf = read_sf(here("data", "oberschulen.gpkg"))
senioren_und_pflegeheime_sf = read_sf(here("data", "senioren_und_pflegeheime.gpkg"))
seniorenbegegnunsstaetten_sf = read_sf(here("data", "seniorenbegegnunsstaetten.gpkg"))

# Data Preparation
# setdiff(stadtteile_sf$blocknr %>% unique(), bevoelkerung_0_bis_5_jahre_sf$stadtteil %>% unique()) # Check for missing quarters: 33, 34, 44
# For better merging
stadtteile_sf$bez = str_replace_all(stadtteile_sf$bez, "/.+", "")
einwohnerzahlen_absolut$X1 = str_replace_all(einwohnerzahlen_absolut$X1, "/.+", "")

cityIdToNumeric = function(bevoelkerung) {
  return (bevoelkerung %>% mutate(stadtteil = as.numeric(stadtteil)) %>% arrange(stadtteil)) # Turn chr to numeric
}
stadtteile_sf = stadtteile_sf %>% mutate(blocknr = as.numeric(blocknr)) %>% arrange(blocknr) 
bevoelkerung_0_bis_5_jahre_sf = cityIdToNumeric(bevoelkerung_0_bis_5_jahre_sf)
bevoelkerung_6_bis_13_jahre_sf = cityIdToNumeric(bevoelkerung_6_bis_13_jahre_sf)
bevoelkerung_14_bis_17_jahre_sf = cityIdToNumeric(bevoelkerung_14_bis_17_jahre_sf)
bevoelkerung_18_bis_26_jahre_sf = cityIdToNumeric(bevoelkerung_18_bis_26_jahre_sf)
bevoelkerung_27_bis_59_jahre_sf = cityIdToNumeric(bevoelkerung_27_bis_59_jahre_sf)
bevoelkerung_60_bis_74_jahre_sf = cityIdToNumeric(bevoelkerung_60_bis_74_jahre_sf)
bevoelkerung_ab_75_jahre_sf = cityIdToNumeric(bevoelkerung_ab_75_jahre_sf)

# Centroids (inferred data)
dresden_stadtteile_centroids = st_centroid(stadtteile_sf)# Centroids of each quarter of Dresden
dresden_centroid = st_centroid(st_union(stadtteile_sf)) # Centroid of Dresden
# Geocode Altmarkt as center of the city to measure distance between quarter centroids and city center
altmarkt_sf = geo("Altmarkt, Dresden", verbose = F) %>% st_as_sf(coords = c("long", "lat"), crs = terra::crs(stadtteile_sf))
dresden_stadtteile_distances_to_center = osrm::osrmTable(
  src = dresden_stadtteile_centroids, dst = altmarkt_sf, osrm.profile = "car") # Distance from each quarter center to Altmarkt by car.
stadtteile_sf$duration_to_altmarkt_by_car = dresden_stadtteile_distances_to_center$durations %>% as.numeric() # Add distance to center to quarters

# Absolute population numbers
stadtteile_sf = merge(stadtteile_sf, einwohnerzahlen_absolut %>% rename(einwohner_abs = X2), by.x = "bez", by.y = "X1", all.x = T)

# OSM Data features
# Schools and Kindergartens
stadtteile_sf$grundschulen_count = grundschulen_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$oberschulen_count = oberschulen_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$kindertagesstaetten_count = kindertagesstaetten_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hausnummer)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
# Restaurants
dresden_restaurants_sf = opq("Dresden") %>% add_osm_feature(key = "amenity", value = "restaurant") %>% osmdata_sf()
stadtteile_sf$restaurants_count = dresden_restaurants_sf$osm_points %>% .[!duplicated(paste0(.$`addr:housenumber`, .$`addr:street`)), ] %>% 
  st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums()
# Parks, 322,1776km^2 free, rest park of 327,2883km^2 total, so 5,1107km^2 parks.
dresden_parks_sf = opq("Dresden") %>% add_osm_feature(key = "leisure", value = "park") %>% osmdata_sf()
stadtteile_sf$parks_surf_tot = dresden_parks_sf$osm_polygons %>%
  st_intersects(x = stadtteile_sf, y = ., sparse = T) %>% lapply(function(x) st_area(dresden_parks_sf$osm_polygons[x, ]) %>% sum) %>% unlist()

# Quiet Areas Data (EU)
quiet_areas_rast <- terra::rast(here("data", "quiet_areas_europe_qis.tif")) %>% 
  terra::crop(st_bbox(st_transform(stadtteile_sf, crs = terra::crs(.)))) %>% 
  terra::project(terra::crs(stadtteile_sf))
# Extract mean values of quiet areas (QIS) per quarter (64 quarters) and bind to sf object.
stadtteile_sf = terra::extract(quiet_areas_rast, stadtteile_sf, fun = max, exact = T, bind = T) %>%  # Set exact = TRUE for exact fraction of cell
  tidyterra::rename(quiet_areas_qis_max = quiet_areas_europe_qis) %>% st_as_sf()
stadtteile_sf = terra::extract(quiet_areas_rast, stadtteile_sf, fun = mean, exact = T, bind = T) %>% 
  rename(quiet_areas_qis_mean = quiet_areas_europe_qis) %>% st_as_sf()
# stadtteile_sf = terra::extract(quiet_areas_rast, stadtteile_sf, fun = min, exact = T, bind = T) %>% 
#   rename(quiet_areas_qis_min = quiet_areas_europe_qis) %>% st_as_sf() # is always zero/nan

# Create bins for population age groups since the age group 27 to 59 years has very high values
bevoelkerung_ab_60_jahre_sf = bevoelkerung_60_bis_74_jahre_sf %>% # upper bin
  mutate(prozent = prozent + bevoelkerung_ab_75_jahre_sf$prozent, 
         anteil = anteil + bevoelkerung_ab_75_jahre_sf$anteil)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_ab_60_jahre_sf)[c("stadtteil", "prozent")] %>% # Merge bevoelkerung_ab_60_jahre_sf into stadtteile_sf
        rename(p_from_60_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all = T)
bevoelkerung_0_bis_26_jahre_sf = bevoelkerung_0_bis_5_jahre_sf %>% # lower bin
  mutate(prozent = prozent + bevoelkerung_6_bis_13_jahre_sf$prozent + bevoelkerung_14_bis_17_jahre_sf$prozent + bevoelkerung_18_bis_26_jahre_sf$prozent, 
         anteil = anteil + bevoelkerung_6_bis_13_jahre_sf$anteil + bevoelkerung_14_bis_17_jahre_sf$anteil + bevoelkerung_18_bis_26_jahre_sf$anteil)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_0_bis_26_jahre_sf)[c("stadtteil", "prozent")] %>% # Merge bevoelkerung_0_bis_26_jahre_sf into stadtteile_sf
        rename(p_0_bis_26_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_27_bis_59_jahre_sf)[c("stadtteil", "prozent")] %>% # Merge bevoelkerung_27_bis_59_jahre_sf into stadtteile_sf
        rename(p_27_to_59_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all = T)
# Merge other features into stadtteile_sf
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_0_bis_5_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_0_to_5_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_6_bis_13_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_6_to_13_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_14_bis_17_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_14_to_17_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_18_bis_26_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_18_to_26_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T) 
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_60_bis_74_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_60_to_74_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_ab_75_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_from_75_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T) 

# Gesundheitsdienste Prep
stadtteile_sf$gesundheitseinrichtungen_count = gesundheitseinrichtungen_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$ambulante_mobile_dienste_count = ambulante_mobile_dienste_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$senioren_pflegeheime_count = senioren_und_pflegeheime_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$seniorenbegegnunsstaetten_count = seniorenbegegnunsstaetten_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$betreutes_wohnen_count = betreutes_wohnen_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$haltestellen_count = haltestellen_sf %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() 
stadtteile_sf$bahnhoefe_count = bahnhoefe_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates

```

# Questions and Hypothesis

This project seeks to explore the reasons for elderly people to live in or move to a certain quarter in Dresden. I think that there are certain factors that make a quarter more attractive for elderly people to live in. These factors could be the availability of health care services, the number of senior citizen meeting places, the number of assisted living facilities, the number of senior and nursing homes, the noise level, and the number of children in the quarter. I hypothesize that these values (see below for more detail on all the values) have an impact on the attractiveness of quarters within Dresden for elderly people to live in. I'll look into these variables in more detail in the following section. The unit of observation is the quarter in Dresden, so I aggregate and/or split the data accordingly. First, I look at the data available and show common statistics, then I'll present some summary statistics, after which I will test the hypothesis and finally discuss the results.

# Data Structure

My base datasets on the age of the population in Dresden as well as data on the quarters are downloaded from the OpenData portal Dresden (<https://opendata.dresden.de>) and are available as vector data (they're already downloaded and included in the data directory of this project). The data includes the population at the place of main residence of Dresden split into different age groups, with each group having a different percentage of the population in each quarter. The population at the place of the main residence includes those people who have their sole residence or their main residence in the area concerned.[^1] The data is available for the year 2021. Since the data on the age of the population is aggregated by quarter, let's take a look at the `r nrow(stadtteile_sf)` different quarters first. In the appendix, you can find a map of the block numbers for each quarter and the corresponding name of the quarter. The blue dot in the following image shows a central meeting point within the center of Dresden, "Altmarkt". This will become relevant later on.

[^1]: For more information on the data, see <https://kommisdd.dresden.de/net3/public/ogc.ashx?Service=Ikx&RenderHint=TargetHtml&NODEID=378&>.

```{r base_map_dresden, echo = F, fig.width = 8}
ggplot(data = stadtteile_sf) +
  geom_sf() + theme_minimal() + 
  # geom_sf(data = dresden_stadtteile_centroids, size = 0.5, colour = "black", alpha = 0.3) + 
  geom_sf_text(aes(label = blocknr), size = 2) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
  geom_sf(data = altmarkt_sf, size = 1, colour = "blue")
```

Another source is the absolute number of people per quarter in Dresden which has been taken from Statista [@dresdenEinwohnerzahlEinzelnenStadtteilen2023] and is part of this repository; it's available as CSV.

The raster data I'm using is a TIF file from the European Environment Agency on the QSI, the quietness suitability ratio. This data is over 2GiB, so you can find it split into pieces in the repository. It has to be merged before usage; it can also be downloaded from the original source[@europeanenvironmentagencyQuietnessSuitabilityIndex2016].

I'm also working with data from OSM (OpenStreetMap) on the number of restaurants and surface area of parks in Dresden per quarter. The data has been downloaded from the OSM API using the `osmdata` package.

Let's now take a look at the data on the age of the population in Dresden. Yellow indicates a higher percentage of the population in the respective age group in the quarter (and blue a lower percentage). The data is available for the following age groups: zero to five years, six to 13 years, 14 to 17 years, 18 to 26 years, 27 to 59 years, 60 to 74 years, and from 75 years. There is no data for three of the 64 quarters in Dresden, namely "Flughafen", "Hellerberge" and "Dresdner Heide"; I'll thus exclude these in the analysis.

Below, you will find two line graphs and two bar charts; the second row shows the binned versions, that is, I merged the age groups because the middle age group from 27 to 59 makes up quite a large percentage compared to the others.

```{r dresden pop maps, echo = F, fig.height=6}
plot_titles = c("0 to 5 years", "6 to 13 years", "14 to 17 years", "18 to 26 years", 
           "27 to 59 years", "60 to 74 years", "From 75 years")
names(plot_titles) = as.character(1:length(plot_titles))
do.call(grid.arrange, 
        c(imap(list(bevoelkerung_0_bis_5_jahre_sf, bevoelkerung_6_bis_13_jahre_sf,  
                    bevoelkerung_14_bis_17_jahre_sf, bevoelkerung_18_bis_26_jahre_sf, 
                    bevoelkerung_27_bis_59_jahre_sf, bevoelkerung_60_bis_74_jahre_sf, 
                    bevoelkerung_ab_75_jahre_sf), 
            \(sf_d, idx) ggplot() + 
              geom_sf(data = stadtteile_sf) + 
              geom_sf(data = sf_d, aes(fill = prozent, color = prozent)) +
              theme_minimal() + 
              ggtitle(plot_titles[idx]) +
              theme(plot.title = element_text(hjust = 0.5, size = 10), axis.text.x = element_blank(), 
                    legend.title = element_text(size = 8)) + 
              scale_fill_viridis_c() +
              scale_colour_viridis_c() + 
              labs(fill = "Percentage", colour = "Percentage")
              ), 
          ncol = 2, nrow = 4, padding = 0))


```

```{r dresden pop stats, echo = F, fig.height = 6}
combined_bev_sf = dplyr::bind_rows(list(bevoelkerung_0_bis_5_jahre_sf, bevoelkerung_6_bis_13_jahre_sf,  
                                        bevoelkerung_14_bis_17_jahre_sf, bevoelkerung_18_bis_26_jahre_sf, 
                                        bevoelkerung_27_bis_59_jahre_sf, bevoelkerung_60_bis_74_jahre_sf, 
                                        bevoelkerung_ab_75_jahre_sf), 
                       .id = "gruppe") 
combined_bev_sf$gruppe = plyr::revalue(as.factor(combined_bev_sf$gruppe), plot_titles)

plot_line_pop_dd = ggplot() + 
  geom_line(data = combined_bev_sf, 
             mapping = aes(x = stadtteil, y = prozent, group = gruppe, col = gruppe), linewidth = 0.3) +
  geom_point(data = combined_bev_sf, 
             mapping = aes(x = stadtteil, y = prozent, group = gruppe, col = gruppe), size = 0.3) +
  theme_minimal() + 
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8), 
        legend.title = element_text(size = 9), axis.text.x = element_blank(), legend.text = element_text(size = 8)) + 
  labs(title = "Population in Dresden by age groups and quarter, \nmeasured in 2021", 
       x = "Quarter Nr.", y = "Percentage", col = "Age Group")

combined_bev_summ = combined_bev_sf %>% 
  group_by(gruppe) %>% 
  summarise(mean = mean(prozent), sd = sd(prozent), median = median(prozent), min = min(prozent), max = max(prozent)) 

plot_bar_pop_dd = ggplot(combined_bev_summ, aes(x = gruppe, y = mean)) + geom_bar(aes(fill = gruppe), stat = "identity", show.legend = F) +
  theme_minimal() + theme(axis.text.x = element_blank()) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),colour = "darkred", width = 0.5, show.legend = F) +
  geom_errorbar(aes(ymin = min, ymax = max), width = 0.5) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8), 
        legend.title = element_text(size = 9)) + 
  labs(title = "Mean percentage of population in \nDresden by age group",
       x = "Age Group", y = "Mean Percentage")


binned_bev_sf = combined_bev_sf
levels(binned_bev_sf$gruppe) = list("0 to 26 years" = c("0 to 5 years", "6 to 13 years", "14 to 17 years", "18 to 26 years"), 
                                                   "27 to 59 years" = "27 to 59 years", 
                                                   "From 60 years" = c("60 to 74 years", "From 75 years"))
binned_bev_sf = aggregate(prozent ~ id + gruppe, binned_bev_sf, sum) 

plot_line_pop_dd_bin = ggplot() + 
  geom_line(data = binned_bev_sf, 
             mapping = aes(x = id, y = prozent, group = gruppe, col = gruppe)) +
  geom_point(data = binned_bev_sf, 
             mapping = aes(x = id, y = prozent, group = gruppe, col = gruppe), size = 0.5) +
  theme_minimal() + 
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8), 
        legend.title = element_text(size = 9), axis.text.x = element_blank(), legend.text = element_text(size = 8)) + 
  labs(title = "Population in Dresden by age groups and quarter, \nmeasured in 2021 (binned)", 
       x = "Quarter Nr.", y = "Percentage", col = "Age Group")


binned_bev_summ = binned_bev_sf %>% 
  group_by(gruppe) %>% 
  summarise(mean = mean(prozent), sd = sd(prozent), median = median(prozent), min = min(prozent), max = max(prozent)) 

plot_bar_pop_dd_bin = ggplot(binned_bev_summ, aes(x = gruppe, y = mean)) + geom_bar(aes(fill = gruppe), stat = "identity", show.legend = F) +
  theme_minimal() + theme(axis.text.x = element_blank()) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),colour = "darkred", width = 0.5, show.legend = F) +
  geom_errorbar(aes(ymin = min, ymax = max), width = 0.5) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8), 
        legend.title = element_text(size = 9)) + 
  labs(title = "Mean percentage of population in \nDresden by age group (binned)",
       x = "Age Group", y = "Mean Percentage")


plot_grid(plot_line_pop_dd, plot_line_pop_dd_bin, plot_bar_pop_dd, plot_bar_pop_dd_bin, ncol = 2, rel_widths = c(3, 2), byrow = F)
```

The biggest percentage of the population is always taken up by the people between 27 to 59 years, with one exception: in "Südvorstadt-Ost" (quarter id 82), the percentage of people between 18 and 26 years is slightly higher (37.4%) than the percentage of people between 27 and 59 years (35.2%). However, this is an outlier of the share of people between 18 and 26 years as their median percentage is 9.2% and their mean 10.84% with a standard deviation of 5.87%; the people between 27 and 59 years have the highest share with a mean of 43.63%, a median of 42.59% and a standard deviation of 5.13%.

People between 27 and 59 years seem to live rather centered (f.ex. 60.19% in "Neustadt"). People in the age group of 60 to 74 years seem to move to outer areas, although it consequently does not seem intuitive why the age group above 75 years mainly lives in inner quarters (block nr. 7, "Johannstadt-Süd") of Dresden. This could be because people in this highest age group prefer more urban, central areas due to various reasons like access to health care or shorter distances. I'll try to look at some more variables for more explanation.

# Statistics Summary

Below, you will find a table of common statistics (mean, median, maximum, minimum and standard deviation) of the used values:

```{r common stats, echo = F}
stadtteile_sf %>% as.data.frame %>% select_if(is.numeric) %>% dplyr::select(!blocknr) %>% summary()
```
It 


[@R-OpenStreetMap]

[@europeanenvironmentagencyQuietnessSuitabilityIndex2016]

It is questionable if all of these variables are useful for the analysis as some of them have very few observations like f.ex. the number of health care facilities. I'll look at the correlation matrix of the variables to see if there are  variables that are highly/lowly correlated and could be in- or excluded from the analysis. 

# Analysis

## Heatmaps

```{r corr dresden pop, echo = F, fig.height = 5}
corr_mat_pop = data.frame(bevoelkerung_0_bis_5_jahre_sf$prozent, bevoelkerung_6_bis_13_jahre_sf$prozent, bevoelkerung_14_bis_17_jahre_sf$prozent, bevoelkerung_18_bis_26_jahre_sf$prozent, bevoelkerung_27_bis_59_jahre_sf$prozent, bevoelkerung_60_bis_74_jahre_sf$prozent, bevoelkerung_ab_75_jahre_sf$prozent)
colnames(corr_mat_pop) = plot_titles
corr_mat_pop = cor(corr_mat_pop) %>% melt()
ggplot(corr_mat_pop, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() + 
   geom_text(aes(label = round(value, 2)), colour = "white") +
  scale_fill_viridis_c() + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(angle = 0)) + 
  labs(x = "", y = "", title = "Correlation Matrix of Dresden Population by Age Group", fill = "Correlation") 

# ggplot(combined_bev_sf) + 
#   geom_point(aes(x = rep(bevoelkerung_ab_75_jahre_sf$prozent, 7), y = combined_bev_sf$prozent[seq_len(nrow(combined_bev_sf)/7*7)], fill = "gruppe")) 
```

Highest positive correlation: 27 to 59 years with 0 to 5 years (72%) and 14 to 17 years with 6 to 13 years (89%). The latter could be explained that in Germany, people usually stay in school at least until the age of 16 years. Also, as children between the age of 0 and 5 usually grow up with their parents at their parent's place, these values should be correlated. All correlations with the highest (oldest) age group are negative except for the correlation with the age group 60 to 74 years. This could be an indicator that elderly people prefer to live in quarters with fewer children and fewer younger people (although it is natural that there is some "squeezing out" since I'm dealing with percentages, so if one group grows, some others have to shrink).

```{r analysis - correlation matrix, echo = F, fig.height = 10}
stadtteile_nums_df = stadtteile_sf %>% as.data.frame %>% select_if(is.numeric) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% dplyr::select(-c(blocknr))
stadtteile_nums_df %>% cor(use="na.or.complete") %>% melt() %>% 
  ggplot(aes(x = Var1, y = Var2, fill = value)) + geom_tile() + scale_fill_viridis_c() + 
  geom_text(aes(label=round(value, 2)), colour = "white", size = 2) +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, hjust = 1), 
                          axis.text.y = element_blank(), legend.position = "bottom") + 
  labs(x = "", y = "", title = "Correlation Matrix of Dresden Quarter Features", fill = "Correlation") 

```
This correlation matrix of all features/variables shows that 

## PCA Analysis

```{r analysis - PCA, echo = F, warnings = F}
pca_stadtteile = prcomp(stadtteile_nums_df, scale = T)
pca_stadtteile$eigenvalues = pca_stadtteile$sdev^2
# biplot(pca_stadtteile)
# PCA(stadtteile_nums_df)
pca_importance <- function(x) {
  vars <- x$sdev^2
  vars <- vars/sum(vars)
  rbind(`Standard deviation` = x$sdev, `Proportion of Variance` = vars, 
      `Cumulative Proportion` = cumsum(vars))
}
pca_importance(pca_stadtteile) %>% kable(format = "latex", booktabs = T, label = "tables", 
                                          caption = "PCA Importance of Dresden Quarter Features")
pca_stadtteile$rotation

#summary(pca_stadtteile)
```

```{r pca screeplot, echo = F, fig.width = 10}
ggplot() + geom_point(aes(x = seq_along(pca_stadtteile$eigenvalues), 
                          y = pca_stadtteile$eigenvalues / sum(pca_stadtteile$eigenvalues))) + 
  geom_line(aes(x = seq_along(pca_stadtteile$eigenvalues), 
                y = pca_stadtteile$eigenvalues / sum(pca_stadtteile$eigenvalues)), linewidth = 0.2) +
  geom_vline(xintercept = 6, colour = "red", linewidth = 0.2) +
  geom_vline(xintercept = 10, colour = "red", linewidth = 0.2) +
  theme_bw() + labs(x = "Principal Component", y = "Proportion of Variance Explained")
```

# Conclusion

# References

::: {#refs}
:::

# Appendix

```{r, echo = F}
stadtteile_df = as.data.frame(stadtteile_sf)[c("blocknr", "bez")] %>% 
  arrange(blocknr)
rows = seq_len(nrow(stadtteile_df) %/% 2)
stadtteile_df$bez = str_replace_all(stadtteile_sf$bez, "/.+", "")
list(stadtteile_df[rows, ], stadtteile_df[-rows, ]) %>%
  kable(format = "latex", booktabs = T, label = "tables", row.names = F,
        caption = "Block numbers and names of quarters in Dresden", 
        col.names = c("Block Number", "Quarter Name")) 
```
