---
title: "Uni GIS Project"
author: "Julius Schmidt"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    toc: true
    fig_caption: true
    number_sections: true
  html_document: default
papersize: a4
bibliography: [references.bib, packages.bib]
---

```{r setup, include = F, warning = F}
knitr::opts_chunk$set(warning = F, message = F)
# Install necessary dependencies from renv.lock
renv::restore()
packages = c(
  "knitr",
  "kableExtra",
  "sf", 
  "tidyverse",
  "here",
  "terra",
  "gridExtra",
  "cowplot",
  "reshape2",
  "magrittr",
  "tidygeocoder",
  "maptiles",
  "osmdata",
  "ncdf4",
  "exactextractr",
  "tidyterra",
  "osrm",
  "factoextra",
  "MASS",
  "FactoMineR",
  "ggfortify"
)
lapply(packages, library, character.only = T)
#knitr::write_bib(file = 'packages.bib')
```

<!-- The report itself should be an R markdown or a Quarto R document of roughly 5 pages (counting without code but including meaningful images or tables that you create), and I should find inside of it a part where you explain the data structure and show summary statistics, a part where you present your research questions and your starting hypothesis, a part where you explain your chosen method of analysis and then apply it and finally a concluding part where you discuss what you find in relation to your hypothesis. -->

```{r Data Setup, echo=F, warning=F}
# Open Data Dresden Portal Data
stadtteile_sf = read_sf(here("data", "stadtteile.gpkg"))
bevoelkerung_0_bis_5_jahre_sf = read_sf(here("data", "bevoelkerung_0_bis_5_jahre.gpkg"))
bevoelkerung_6_bis_13_jahre_sf = read_sf(here("data", "bevoelkerung_6_bis_13_jahre.gpkg"))
bevoelkerung_14_bis_17_jahre_sf = read_sf(here("data", "bevoelkerung_14_bis_17_jahre.gpkg"))
bevoelkerung_18_bis_26_jahre_sf = read_sf(here("data", "bevoelkerung_18_bis_26_jahre.gpkg"))
bevoelkerung_27_bis_59_jahre_sf = read_sf(here("data", "bevoelkerung_27_bis_59_jahre.gpkg"))
bevoelkerung_60_bis_74_jahre_sf = read_sf(here("data", "bevoelkerung_60_bis_74_jahre.gpkg"))
bevoelkerung_ab_75_jahre_sf = read_sf(here("data", "bevoelkerung_ab_75_jahre.gpkg"))
ambulante_mobile_dienste_sf = read_sf(here("data",  "ambulante_mobile_dienste.gpkg"))
bahnhoefe_sf = read_sf(here("data", "bahnhoefe_dresden.gpkg"))
betreutes_wohnen_sf = read_sf(here("data", "betreutes_wohnen.gpkg"))
einwohnerzahlen_absolut = read_csv2(here("data", "einwohnerzahlen_absolut.csv"), col_names = F, show_col_types = F)
gesundheitseinrichtungen_sf = read_sf(here("data", "gesundheitseinrichtungen.gpkg"))
grundschulen_sf = read_sf(here("data", "grundschulen.gpkg"))
haltestellen_sf = read_sf(here("data", "haltestellen_dresden.gpkg"))
kindertagesstaetten_sf = read_sf(here("data", "kindertagesstaetten.gpkg"))
oberschulen_sf = read_sf(here("data", "oberschulen.gpkg"))
senioren_und_pflegeheime_sf = read_sf(here("data", "senioren_und_pflegeheime.gpkg"))
seniorenbegegnunsstaetten_sf = read_sf(here("data", "seniorenbegegnunsstaetten.gpkg"))

# Data Preparation
# setdiff(stadtteile_sf$blocknr %>% unique(), bevoelkerung_0_bis_5_jahre_sf$stadtteil %>% unique()) # Check for missing quarters: 33, 34, 44
# For better merging
stadtteile_sf$bez = str_replace_all(stadtteile_sf$bez, "/.+", "")
einwohnerzahlen_absolut$X1 = str_replace_all(einwohnerzahlen_absolut$X1, "/.+", "")

cityIdToNumeric = function(bevoelkerung) {
  return (bevoelkerung %>% mutate(stadtteil = as.numeric(stadtteil)) %>% arrange(stadtteil)) # Turn chr to numeric
}
stadtteile_sf = stadtteile_sf %>% mutate(blocknr = as.numeric(blocknr)) %>% arrange(blocknr) 
bevoelkerung_0_bis_5_jahre_sf = cityIdToNumeric(bevoelkerung_0_bis_5_jahre_sf)
bevoelkerung_6_bis_13_jahre_sf = cityIdToNumeric(bevoelkerung_6_bis_13_jahre_sf)
bevoelkerung_14_bis_17_jahre_sf = cityIdToNumeric(bevoelkerung_14_bis_17_jahre_sf)
bevoelkerung_18_bis_26_jahre_sf = cityIdToNumeric(bevoelkerung_18_bis_26_jahre_sf)
bevoelkerung_27_bis_59_jahre_sf = cityIdToNumeric(bevoelkerung_27_bis_59_jahre_sf)
bevoelkerung_60_bis_74_jahre_sf = cityIdToNumeric(bevoelkerung_60_bis_74_jahre_sf)
bevoelkerung_ab_75_jahre_sf = cityIdToNumeric(bevoelkerung_ab_75_jahre_sf)

# Centroids (inferred data)
dresden_stadtteile_centroids = st_centroid(stadtteile_sf)# Centroids of each quarter of Dresden
dresden_centroid = st_centroid(st_union(stadtteile_sf)) # Centroid of Dresden
# Geocode Altmarkt as center of the city to measure distance between quarter centroids and city center
altmarkt_sf = geo("Altmarkt, Dresden", verbose = F) %>% st_as_sf(coords = c("long", "lat"), crs = terra::crs(stadtteile_sf))
dresden_stadtteile_distances_to_center = osrm::osrmTable(
  src = dresden_stadtteile_centroids, dst = altmarkt_sf, osrm.profile = "car") # Distance from each quarter center to Altmarkt by car.
stadtteile_sf$duration_to_altmarkt_by_car = dresden_stadtteile_distances_to_center$durations %>% as.numeric() # Add distance to center to quarters

# Absolute population numbers
stadtteile_sf = merge(stadtteile_sf, einwohnerzahlen_absolut %>% rename(einwohner_abs = X2), by.x = "bez", by.y = "X1", all.x = T)

# OSM Data features
# Schools and Kindergartens
stadtteile_sf$grundschulen_count = grundschulen_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$oberschulen_count = oberschulen_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$kindertagesstaetten_count = kindertagesstaetten_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hausnummer)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
# Restaurants
dresden_restaurants_sf = opq("Dresden") %>% add_osm_feature(key = "amenity", value = "restaurant") %>% osmdata_sf()
stadtteile_sf$restaurants_count = dresden_restaurants_sf$osm_points %>% .[!duplicated(paste0(.$`addr:housenumber`, .$`addr:street`)), ] %>% 
  st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums()
# Parks, 322,1776km^2 free, rest park of 327,2883km^2 total, so 5,1107km^2 parks.
dresden_parks_sf = opq("Dresden") %>% add_osm_feature(key = "leisure", value = "park") %>% osmdata_sf()
stadtteile_sf$parks_surf_tot = dresden_parks_sf$osm_polygons %>%
  st_intersects(x = stadtteile_sf, y = ., sparse = T) %>% lapply(function(x) st_area(dresden_parks_sf$osm_polygons[x, ]) %>% sum) %>% unlist()

# Quiet Areas Data (EU)
quiet_areas_rast <- terra::rast(here("data", "quiet_areas_europe_qis.tif")) %>% 
  terra::crop(st_bbox(st_transform(stadtteile_sf, crs = terra::crs(.)))) %>% 
  terra::project(terra::crs(stadtteile_sf))
# Extract mean values of quiet areas (QIS) per quarter (64 quarters) and bind to sf object.
stadtteile_sf = terra::extract(quiet_areas_rast, stadtteile_sf, fun = max, exact = T, bind = T) %>%  # Set exact = TRUE for exact fraction of cell
  tidyterra::rename(quiet_areas_qis_max = quiet_areas_europe_qis) %>% st_as_sf()
stadtteile_sf = terra::extract(quiet_areas_rast, stadtteile_sf, fun = mean, exact = T, bind = T) %>% 
  rename(quiet_areas_qis_mean = quiet_areas_europe_qis) %>% st_as_sf()
# stadtteile_sf = terra::extract(quiet_areas_rast, stadtteile_sf, fun = min, exact = T, bind = T) %>% 
#   rename(quiet_areas_qis_min = quiet_areas_europe_qis) %>% st_as_sf() # is always zero/nan

# Create bins for population age groups since the age group 27 to 59 years has very high values
bevoelkerung_ab_60_jahre_sf = bevoelkerung_60_bis_74_jahre_sf %>% # upper bin
  mutate(prozent = prozent + bevoelkerung_ab_75_jahre_sf$prozent, 
         anteil = anteil + bevoelkerung_ab_75_jahre_sf$anteil)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_ab_60_jahre_sf)[c("stadtteil", "prozent")] %>% # Merge bevoelkerung_ab_60_jahre_sf into stadtteile_sf
        rename(p_from_60_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all = T)
bevoelkerung_0_bis_26_jahre_sf = bevoelkerung_0_bis_5_jahre_sf %>% # lower bin
  mutate(prozent = prozent + bevoelkerung_6_bis_13_jahre_sf$prozent + bevoelkerung_14_bis_17_jahre_sf$prozent + bevoelkerung_18_bis_26_jahre_sf$prozent, 
         anteil = anteil + bevoelkerung_6_bis_13_jahre_sf$anteil + bevoelkerung_14_bis_17_jahre_sf$anteil + bevoelkerung_18_bis_26_jahre_sf$anteil)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_0_bis_26_jahre_sf)[c("stadtteil", "prozent")] %>% # Merge bevoelkerung_0_bis_26_jahre_sf into stadtteile_sf
        rename(p_0_bis_26_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_27_bis_59_jahre_sf)[c("stadtteil", "prozent")] %>% # Merge bevoelkerung_27_bis_59_jahre_sf into stadtteile_sf
        rename(p_27_to_59_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all = T)
# Merge other features into stadtteile_sf
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_0_bis_5_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_0_to_5_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_6_bis_13_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_6_to_13_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_14_bis_17_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_14_to_17_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_18_bis_26_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_18_to_26_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T) 
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_60_bis_74_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_60_to_74_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T)
stadtteile_sf = merge(stadtteile_sf, as.data.frame(bevoelkerung_ab_75_jahre_sf)[c("stadtteil", "prozent")] %>% 
        rename(p_from_75_yrs = prozent), by.x = "blocknr", by.y = "stadtteil", all.x = T) 

# Gesundheitsdienste Prep
stadtteile_sf$gesundheitseinrichtungen_count = gesundheitseinrichtungen_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$ambulante_mobile_dienste_count = ambulante_mobile_dienste_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$senioren_pflegeheime_count = senioren_und_pflegeheime_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$seniorenbegegnunsstaetten_count = seniorenbegegnunsstaetten_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$betreutes_wohnen_count = betreutes_wohnen_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates
stadtteile_sf$haltestellen_count = haltestellen_sf %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() 
stadtteile_sf$bahnhoefe_count = bahnhoefe_sf %>% 
  .[!duplicated(paste0(.$strasse, .$hnr)), ] %>% st_intersects(x = stadtteile_sf, y = ., sparse = F) %>% rowSums() # filter out duplicates

```

# Questions and Hypothesis

This project seeks to explore the reasons for elderly people to live in or move to a certain quarter in Dresden. I think that there are certain factors that make a quarter more attractive for elderly people to live in. These factors could be the availability of health care services, the number of senior citizen meeting places, the number of assisted living facilities, the number of senior and nursing homes, the noise level, and the number of children in the quarter. I hypothesize that these values (see below for more detail on all the values) have an impact on the attractiveness of quarters within Dresden for elderly people to live in. I'll look into these variables in more detail in the following section. The unit of observation is the quarter in Dresden, so I aggregate and/or split the data accordingly. First, I look at the data available and show common statistics, then I'll present some summary statistics, after which I will test the hypothesis and finally discuss the results.

# Data Structure

My base datasets on the age of the population in Dresden as well as data on the quarters are downloaded from the OpenData portal Dresden (<https://opendata.dresden.de>) and are available as vector data (they're already downloaded and included in the data directory of this project). The data includes the population at the place of main residence of Dresden split into different age groups, with each group having a different percentage of the population in each quarter. The population at the place of the main residence includes those people who have their sole residence or their main residence in the area concerned.[^1] The data is available for the year 2021. Since the data on the age of the population is aggregated by quarter, let's take a look at the `r nrow(stadtteile_sf)` different quarters first. In the appendix, you can find a map of the block numbers for each quarter and the corresponding name of the quarter. The blue dot in the following image shows a central meeting point within the center of Dresden, "Altmarkt". This will become relevant later on.

[^1]: For more information on the data, see <https://kommisdd.dresden.de/net3/public/ogc.ashx?Service=Ikx&RenderHint=TargetHtml&NODEID=378&>.

```{r base_map_dresden, echo = F, fig.width = 8}
ggplot(data = stadtteile_sf) +
  geom_sf() + theme_minimal() + 
  # geom_sf(data = dresden_stadtteile_centroids, size = 0.5, colour = "black", alpha = 0.3) + 
  geom_sf_text(aes(label = blocknr), size = 2) +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()) +
  geom_sf(data = altmarkt_sf, size = 1, colour = "blue")
```

Another source is the absolute number of people per quarter in Dresden which has been taken from Statista [@dresdenEinwohnerzahlEinzelnenStadtteilen2023] and is part of this repository; it's available as CSV.

The raster data I'm using is a TIF file from the European Environment Agency on the QSI, the quietness suitability ratio. This data is over 2GiB, so you can find it split into pieces in the repository. It has to be merged before usage; it can also be downloaded from the original source[@europeanenvironmentagencyQuietnessSuitabilityIndex2016][^2].

[^2]: 2016 Quietness Suitability Index (QSI) raster data from the European Environment Agency, available at <https://sdi.eea.europa.eu/data/e9151c34-da65-48b9-a2ca-b9b835480812>.

I'm also working with data from OSM (OpenStreetMap) on the number of restaurants and surface area of parks in Dresden per quarter. The data has been downloaded from the OSM API using the `osmdata` package.

Let's now take a look at the data on the age of the population in Dresden. Yellow indicates a higher percentage of the population in the respective age group in the quarter (and blue a lower percentage). The data is available for the following age groups: zero to five years, six to 13 years, 14 to 17 years, 18 to 26 years, 27 to 59 years, 60 to 74 years, and from 75 years. There is no data for three of the 64 quarters in Dresden, namely "Flughafen", "Hellerberge" and "Dresdner Heide"; I'll thus exclude these in the analysis.

Below, you will find two line graphs and two bar charts; the second row shows the binned versions, that is, I merged the age groups because the middle age group from 27 to 59 makes up quite a large percentage compared to the others.

```{r dresden pop maps, echo = F, fig.height=6}
plot_titles = c("0 to 5 years", "6 to 13 years", "14 to 17 years", "18 to 26 years", 
           "27 to 59 years", "60 to 74 years", "From 75 years")
names(plot_titles) = as.character(1:length(plot_titles))
do.call(grid.arrange, 
        c(imap(list(bevoelkerung_0_bis_5_jahre_sf, bevoelkerung_6_bis_13_jahre_sf,  
                    bevoelkerung_14_bis_17_jahre_sf, bevoelkerung_18_bis_26_jahre_sf, 
                    bevoelkerung_27_bis_59_jahre_sf, bevoelkerung_60_bis_74_jahre_sf, 
                    bevoelkerung_ab_75_jahre_sf), 
            \(sf_d, idx) ggplot() + 
              geom_sf(data = stadtteile_sf) + 
              geom_sf(data = sf_d, aes(fill = prozent, color = prozent)) +
              theme_minimal() + 
              ggtitle(plot_titles[idx]) +
              theme(plot.title = element_text(hjust = 0.5, size = 10), axis.text.x = element_blank(), 
                    legend.title = element_text(size = 8)) + 
              scale_fill_viridis_c() +
              scale_colour_viridis_c() + 
              labs(fill = "Percentage", colour = "Percentage")
              ), 
          ncol = 2, nrow = 4, padding = 0))


```

```{r dresden pop stats, echo = F, fig.height = 6}
combined_bev_sf = dplyr::bind_rows(list(bevoelkerung_0_bis_5_jahre_sf, bevoelkerung_6_bis_13_jahre_sf,  
                                        bevoelkerung_14_bis_17_jahre_sf, bevoelkerung_18_bis_26_jahre_sf, 
                                        bevoelkerung_27_bis_59_jahre_sf, bevoelkerung_60_bis_74_jahre_sf, 
                                        bevoelkerung_ab_75_jahre_sf), 
                       .id = "gruppe") 
combined_bev_sf$gruppe = plyr::revalue(as.factor(combined_bev_sf$gruppe), plot_titles)

plot_line_pop_dd = ggplot() + 
  geom_line(data = combined_bev_sf, 
             mapping = aes(x = stadtteil, y = prozent, group = gruppe, col = gruppe), linewidth = 0.3) +
  geom_point(data = combined_bev_sf, 
             mapping = aes(x = stadtteil, y = prozent, group = gruppe, col = gruppe), size = 0.3) +
  theme_minimal() + 
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8), 
        legend.title = element_text(size = 9), axis.text.x = element_blank(), legend.text = element_text(size = 8)) + 
  labs(title = "Population in Dresden by age groups and quarter, \nmeasured in 2021", 
       x = "Quarter Nr.", y = "Percentage", col = "Age Group")

combined_bev_summ = combined_bev_sf %>% 
  group_by(gruppe) %>% 
  summarise(mean = mean(prozent), sd = sd(prozent), median = median(prozent), min = min(prozent), max = max(prozent)) 

plot_bar_pop_dd = ggplot(combined_bev_summ, aes(x = gruppe, y = mean)) + geom_bar(aes(fill = gruppe), stat = "identity", show.legend = F) +
  theme_minimal() + theme(axis.text.x = element_blank()) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),colour = "darkred", width = 0.5, show.legend = F) +
  geom_errorbar(aes(ymin = min, ymax = max), width = 0.5) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8), 
        legend.title = element_text(size = 9)) + 
  labs(title = "Mean percentage of population in \nDresden by age group",
       x = "Age Group", y = "Mean Percentage")


binned_bev_sf = combined_bev_sf
levels(binned_bev_sf$gruppe) = list("0 to 26 years" = c("0 to 5 years", "6 to 13 years", "14 to 17 years", "18 to 26 years"), 
                                                   "27 to 59 years" = "27 to 59 years", 
                                                   "From 60 years" = c("60 to 74 years", "From 75 years"))
binned_bev_sf = aggregate(prozent ~ id + gruppe, binned_bev_sf, sum) 

plot_line_pop_dd_bin = ggplot() + 
  geom_line(data = binned_bev_sf, 
             mapping = aes(x = id, y = prozent, group = gruppe, col = gruppe)) +
  geom_point(data = binned_bev_sf, 
             mapping = aes(x = id, y = prozent, group = gruppe, col = gruppe), size = 0.5) +
  theme_minimal() + 
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8), 
        legend.title = element_text(size = 9), axis.text.x = element_blank(), legend.text = element_text(size = 8)) + 
  labs(title = "Population in Dresden by age groups and quarter, \nmeasured in 2021 (binned)", 
       x = "Quarter Nr.", y = "Percentage", col = "Age Group")


binned_bev_summ = binned_bev_sf %>% 
  group_by(gruppe) %>% 
  summarise(mean = mean(prozent), sd = sd(prozent), median = median(prozent), min = min(prozent), max = max(prozent)) 

plot_bar_pop_dd_bin = ggplot(binned_bev_summ, aes(x = gruppe, y = mean)) + geom_bar(aes(fill = gruppe), stat = "identity", show.legend = F) +
  theme_minimal() + theme(axis.text.x = element_blank()) +
  geom_errorbar(aes(ymin = mean - sd, ymax = mean + sd),colour = "darkred", width = 0.5, show.legend = F) +
  geom_errorbar(aes(ymin = min, ymax = max), width = 0.5) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8), 
        legend.title = element_text(size = 9)) + 
  labs(title = "Mean percentage of population in \nDresden by age group (binned)",
       x = "Age Group", y = "Mean Percentage")


plot_grid(plot_line_pop_dd, plot_line_pop_dd_bin, plot_bar_pop_dd, plot_bar_pop_dd_bin, ncol = 2, rel_widths = c(3, 2), byrow = F)
```

The biggest percentage of the population is always taken up by the people between 27 to 59 years, with one exception: in "Südvorstadt-Ost" (quarter id 82), the percentage of people between 18 and 26 years is slightly higher (37.4%) than the percentage of people between 27 and 59 years (35.2%). However, this is an outlier of the share of people between 18 and 26 years as their median percentage is 9.2% and their mean 10.84% with a standard deviation of 5.87%; the people between 27 and 59 years have the highest share with a mean of 43.63%, a median of 42.59% and a standard deviation of 5.13%.

People between 27 and 59 years seem to live rather centered (f.ex. 60.19% in "Neustadt"). People in the age group of 60 to 74 years seem to move to outer areas, although it consequently does not seem intuitive why the age group above 75 years mainly lives in inner quarters (block nr. 7, "Johannstadt-Süd") of Dresden. This could be because people in this highest age group prefer more urban, central areas due to various reasons like access to health care or shorter distances. I'll try to look at some more variables for more explanation.

# Statistics Summary

Below, you will find a table of common statistics (mean, median, maximum, minimum and standard deviation) of the used values:

```{r common stats, echo = F}
stadtteile_sf %>% as.data.frame %>% select_if(is.numeric) %>% dplyr::select(!blocknr) %>% summary()
```

The inspected variables are the surface area per quarter, the absolute number of inhabitants, the number of primary and secondary schools, the number of kindergartens, the number of restaurants, of parks, of health care facilities, of senior citizen meeting places, of assisted living facilities, of senior and nursing homes, of bus stops and of train stations. Also, the duration in minutes to the center of Dresden (I chose Altmarkt as a popular meeting point) by car is included, which was calculated from the centers of each quarter to Altmarkt and using OSM. The percentage of the population in Dresden by age group is also included, including the three binned versions of the age groups.

It is questionable if all of these variables are useful for the analysis as some of them have very few observations like f.ex. the number of health care facilities. I'll look at the correlation matrix of the variables to see if there are variables that are highly/lowly correlated and could be in- or excluded from the analysis.

# Analysis

## Heatmaps

```{r corr dresden pop, echo = F, fig.height = 5}
corr_mat_pop = data.frame(bevoelkerung_0_bis_5_jahre_sf$prozent, bevoelkerung_6_bis_13_jahre_sf$prozent, bevoelkerung_14_bis_17_jahre_sf$prozent, bevoelkerung_18_bis_26_jahre_sf$prozent, bevoelkerung_27_bis_59_jahre_sf$prozent, bevoelkerung_60_bis_74_jahre_sf$prozent, bevoelkerung_ab_75_jahre_sf$prozent)
colnames(corr_mat_pop) = plot_titles
corr_mat_pop = cor(corr_mat_pop) %>% melt()
ggplot(corr_mat_pop, aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() + 
   geom_text(aes(label = round(value, 2)), colour = "white") +
  scale_fill_viridis_c() + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), axis.text.y = element_text(angle = 0)) + 
  labs(x = "", y = "", title = "Correlation Matrix of Dresden Population by Age Group", fill = "Correlation") 

# ggplot(combined_bev_sf) + 
#   geom_point(aes(x = rep(bevoelkerung_ab_75_jahre_sf$prozent, 7), y = combined_bev_sf$prozent[seq_len(nrow(combined_bev_sf)/7*7)], fill = "gruppe")) 
```

The highest positive correlations are: 27 to 59 years with 0 to 5 years (72%) and 14 to 17 years with 6 to 13 years (89%). The latter could be explained by the fact that in Germany, people usually stay in school at least until the age of 16 years. Also, as children between the age of 0 and 5 usually grow up with their parents at their parent's place, these values should be correlated. All correlations with the highest (oldest) age group are negative except for the correlation with the age group 60 to 74 years. This could be an indicator that elderly people prefer to live in quarters with fewer children and fewer younger people (although it is natural that there is some "squeezing out" since I'm dealing with percentages, so if one group grows, some others inevitably have to shrink).

```{r analysis - correlation matrix, echo = F, fig.height = 10}
stadtteile_nums_df = stadtteile_sf %>% as.data.frame %>% select_if(is.numeric) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% dplyr::select(-c(blocknr))
stadtteile_nums_df %>% cor(use="na.or.complete") %>% melt() %>% 
  ggplot(aes(x = Var1, y = Var2, fill = value)) + geom_tile() + scale_fill_viridis_c() + 
  scale_x_discrete(labels = paste0(names(stadtteile_nums_df), " (", 1:ncol(stadtteile_nums_df), ")")) +
  scale_y_discrete(labels = 1:ncol(stadtteile_nums_df)) +
  geom_text(aes(label=round(value, 2)), colour = "white", size = 2) +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, hjust = 1), 
                          axis.text.y = element_text(), legend.position = "bottom") + 
  labs(x = "", y = "", title = "Correlation Matrix of Dresden Quarter Features", fill = "Correlation") 

```

As the correlation matrix is rather large, I'd like to highlight some of the most interesting correlations: There is a high correlation (82%) between the surface area and the QSI (Quiet Suitability Index) per quarter - this might be because quarters with a larger surface usually also have more free and green space (like f.ex. parks) or because of a lower population density. Whereas the former can't be supported directly by the correlation between the number of parks and the surface area (-7%), the latter can be supported by a negative (-52%) correlation between the absolute number of inhabitants and the surface area. Since larger quarters in Dresden are those that are further away from the center, it would support the argument that elderly people between 60 and 74 years are to move to quieter areas, although it would take them longer to drive into the city center. As we saw above, people from the age of 75 move to the inner quarters again, which could be explained by the fact that they prefer to live in more urban areas with better access to health care facilities and shorter distances to the city center. There is indeed a correlation, although rather weak (16%), to the number of health care facilities, however, for the age group between 60 and 74 years, this correlation is -12%, so this could be a reason why the highest age group moves to the city center again. One can also note that there is a correlation of 15% for the age group of 75+ and a negative correlation of -28% for the age group 60 to 74 years for senior nursing homes, which would again support the thesis of old people moving to town again because of general access to health care (and all that is included, senior nursing homes etc.). However, there is almost no correlation (1% and -4%) for the 75+ group with the number of senior meeting places and the number of assisted living facilities, respectively. There is also positive correlation (18% and 14%), although again rather weak, between the age group 75+ and the number of bus stops and train stations. However, there is a positive correlation between the age group 75+ and primary schools (10%) and kindergartens (18%), which weakens the hypothesis that elderly people prefer to live in quarters with fewer children, although one could argue that this is because they prefer general access to health care, as the age group between 60 and 74 years f.ex. tends to "avoid" people between 18 and 26 years (-31%).

## PCA Analysis

In the following, you will see a result table of a PCA analysis and a screeplot.

```{r analysis - PCA, echo = F, warnings = F}
pca_stadtteile = prcomp(stadtteile_nums_df, scale = T)
pca_stadtteile$eigenvalues = pca_stadtteile$sdev^2
# biplot(pca_stadtteile)
# PCA(stadtteile_nums_df)
pca_importance <- function(x) {
  vars <- x$sdev^2
  vars <- vars/sum(vars)
  rbind(`Stand. Dev.` = x$sdev, `Prop. of Var.` = vars, 
      `Cum Prop.` = cumsum(vars))
}
pca_df = pca_importance(pca_stadtteile) 
# pca_stadtteile$rotation

list(pca_df[, 1:7], pca_df[, 8:14], pca_df[, 15:21], pca_df[, 22:ncol(pca_df)]) %>%
  kable(format = "latex", booktabs = T, label = "tables",
        caption = "PCA Importance of Dresden Quarter Features") %>% 
  kable_styling(latex_options = "hold_position")

#summary(pca_stadtteile)
```

```{r pca screeplot, echo = F, fig.width = 10}
ggplot() + geom_point(aes(x = seq_along(pca_stadtteile$eigenvalues), 
                          y = pca_stadtteile$eigenvalues / sum(pca_stadtteile$eigenvalues))) + 
  geom_line(aes(x = seq_along(pca_stadtteile$eigenvalues), 
                y = pca_stadtteile$eigenvalues / sum(pca_stadtteile$eigenvalues)), linewidth = 0.2) +
  geom_vline(xintercept = 6, colour = "red", linewidth = 0.2) +
  geom_vline(xintercept = 10, colour = "red", linewidth = 0.2) +
  theme_bw() + labs(x = "Principal Component", y = "Proportion of Variance Explained")
```

One can see that the first PC explains about 24% of the variance and the second PC about 17%, after which the percentage halves. The first six components explain about 68% (first red vertical line in graph above) and the first 10 about 83% (second red vertical line in graph above) of the variance. This means that the first six components are the most important ones and should be used for further analysis. I'll use k-Means clustering with four clusters to see if the quarters can be grouped into four different groups, after which I will plot the two first PCs and colour the clusters by k-Means to see if the PCs can explain the clustering.

```{r clustering, echo = F, fig.width = 10}
kmeans = eclust(stadtteile_nums_df, k=4)
```

```{r clustering plot, echo = F, fig.width = 8}
autoplot(pca_stadtteile, data = kmeans, colour = "cluster")
```

One can see that there is some overlap between the clusters and even a few outliers, especially in the second cluster. The first cluster seems to be the most distinct one, although it is not very clear if the PCs can explain the clustering.

# Conclusion

Although I found some correlations between variables that seem intuitive (f.ex. people in the 75+ age group moving the the center of the city to have better access to general health care services), other parts of the initial hypothesis like dependence on number of restaurants or number of children in the quarter could not be supported by the data. The thought that elderly people move to the center of the city because of general health services seems to be backed by the data. The PCA analysis did not show a clear clustering of the quarters based on the features I used. This could be due to the fact that the data is not detailed enough or that the variables I used are not the right ones to explain the attractiveness of quarters for elderly people to live in. Also, the data on the age of the population in Dresden could be more detailed, for example, by splitting the age group 27 to 59 years into smaller groups. Also, regarding the QSI that generally increases in larger areas and consequently in areas further away from the center, people between 60 and 74 years seem to prefer quieter areas, while people at 75+ seem to value access to general health services higher.

# References

::: {#refs}
:::

# Appendix

```{r, echo = F}
stadtteile_df = as.data.frame(stadtteile_sf)[c("blocknr", "bez")] %>% 
  arrange(blocknr)
rows = seq_len(nrow(stadtteile_df) %/% 2)
stadtteile_df$bez = str_replace_all(stadtteile_sf$bez, "/.+", "")
list(stadtteile_df[rows, ], stadtteile_df[-rows, ]) %>%
  kable(format = "latex", booktabs = T, label = "tables", row.names = F,
        caption = "Block numbers and names of quarters in Dresden", 
        col.names = c("Block Number", "Quarter Name")) 
```
